---
title: "Covid19 vaccine study Fall2021"
author: "Rachel Gilpatrick"
date: "11-29-2021"
output: 
  html_document:
    toc: true
    toc_depth: 5

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="fall21.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$: There is no difference in covid-19 vaccine effectiveness among different subgroups (males, females, drug users, LGBTQ)

$H_A$: There is a difference in covid-19 vaccine effectiveness among different subgroups(males, femlaes, drug users, LGBTQ)


# Males



```{r}
dd2 <- guy %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```






This plot represents the population of guys(males) people and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. 


The plot above shows that a shocking 58% of the people that took the vaccine did in fact get covid-19, while only approximately 42% of the people taking the placebo contracted covid-19. A substantially higher percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be ineffective in protection against covid-19. The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical Results




```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of guys(males) that got covid-19 after taking the placebo and taking the vaccine.
Among the guys, 58% got covid-19 after taking the vaccine, while approximately 42% of guys that were given the placebo contracted covid-19. The vaccine is clearly not effective in preventing covid-19 in guys(males) and it is not recommended. Those who took th eplacebo were less likely to contract covid-19.
The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
guyfish <- fisher.test(table1)
guyfish
```




The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 0.0005801. This is an awfully small number. seeing this number means that these results are not likely to occur by chance.



The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sex, is the favorable option.

The odd ratio above illustrates a ratio of 0.68. the guys(men) who received the vaccine were 1.5 times more likely to get covid-19 than those who received the placebo. 


```{r}
effect <- guyfish$p.value/(guyfish$p.value +1)
effectivenessguy = guyfish$estimate/(guyfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guyfish$conf.int[1]/(guyfish$conf.int[1] +1))*100
hi=(guyfish$conf.int[2]/(guyfish$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for guys(males). If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting guys from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the guys can be seen above as 40. The effectiveness confidence interval for the guys is 35.09 to 45.94.

# Females



```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```






This plot represents the population of gals(females) and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. 


The plot above shows that only 33% of the gals that took the vaccine did in fact get covid-19, while 67% of the people taking the placebo contracted covid-19.  A substantially higher percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for the gals (females). The odds of getting covid-19 while taking the vaccine is 1:3. The women that took the drug were half as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of gals that got covid-19 or not.
In the second plot, the one above, The left two bars of the graph are all the gals that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of gals who got covid-19, now represents the percentage of gals that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical Results




```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of gals(females) that got covid-19 after taking the placebo and taking the vaccine.
among the gals, 33% got covid-19 after taking the vaccine, while approximately 67% of the gals that were given the placebo contracted covid-19. The vaccine is clearly effetive in preventing covid-19 in gals(females) and it is not recommended. Those who took the placebo were much less likely to contract covid-19.
The odds of getting covid-19 for gals while taking the vaccine is 1:2. The gals taht took the drug were half as likely to get covid-19 (33%) than those taking the placebo (67%).

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
galfish <- fisher.test(table1)
galfish
```




The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 2e-16. This is an awfully small number. This number means that these results are not likely to occur by chance.



The null hypothesis predicted that the infections rate related to covid-19 vaccination effectiveness are independent of subgroup. Since it is clear that subgroup does play a role in the covid infection rate in the vaccinated individuals, the alternative hypothesis, that there is a correlation between covid-19 vaccination effectiveness and subgroups, is the favorable option.

The odd ratio above illustrates a ratio of 2.00. the gals(females) who received the vaccine were half as likely to get covid-19 than those who received the placebo. 

```{r}
effect <- galfish$p.value/(galfish$p.value +1)
effectivenessgal = galfish$estimate/(galfish$estimate +1)*100
print(effectivenessgal)
```


```{r}
low=(galfish$conf.int[1]/(galfish$conf.int[1] +1))*100
hi=(galfish$conf.int[2]/(galfish$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for gals(females). If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting gals (females) from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the gals (females) can be seen above as 67. The effectiveness confidence interval for the gals is 55.09 to 75.05.

# Drug Users


```{r}
dd2 <- Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```






This plot represents the population of Drug users and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. 


The plot above shows that only 33% of the drug users that took the vaccine did in fact get covid-19, while 67% of the people taking the placebo contracted covid-19.  A substantially lower percentage of people who took tha vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for the gals (females). The odds of getting covid-19 while taking the vaccine is 1:3. The drug users that took the drug were half as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of drug users that got covid-19 or not.
In the second plot, the one above, The left two bars of the graph are all the drug users that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of drug users who got covid-19, now represents the percentage of drug users that got covid-19 after taking the placebo or taking the Vaccine. 


The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected, contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical Results




```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of drug users that got covid-19 after taking the placebo and taking the vaccine. 
Among the guys, 33% got covid-19 after taking the vaccine, while the approximately 67% of drug users that were given the placebo contracted covid-19. The vaccine is clearly effective in preventing covid-19 in drug users and it is recommended. Those who took the placebo were more likely to contract covid-19.
The odds of getting covid-19 while taking the vaccine is 1:2. The drug users that took the drug were half as likely to get covid-19 (33%) than those taking the placebo (67%).


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
Druguserfish <- fisher.test(table1)
Druguserfish
```


The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at .002. This is an awfully small number. seeing this number means that these results are not likely to occur by chance.



The null hypothesis predicted that the civid-19 vaccine effectiveness is independent of subgroup. Since it is clear that subgroup does play a role in the rate of infection in individuals who were vaccinated, the alternative hypothesis, that there is a correlation between covid-19 infection rate in those vaccinated and subgroups, is the favorable option.

The odd ratio above illustrates a ratio of 1.8. The drug users who received the vaccine were half likely to get covid-19 than those who received the placebo. 


```{r}
effect <- Druguserfish$p.value/(Druguserfish$p.value +1)
effectivenessDruguser = Druguserfish$estimate/(Druguserfish$estimate +1)*100
print(effectivenessDruguser)
```

```{r}
low=(Druguserfish$conf.int[1]/(Druguserfish$conf.int[1] +1))*100
hi=(Druguserfish$conf.int[2]/(Druguserfish$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for drug users. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting drug users from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the drug users can be seen above as 65. The effectiveness confidence interval for the drug users is 55.09 to 75.05.

# LGBTQ

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```






This plot represents the population of drug users and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo.


The plot above shows that only 33% of the LGBTQ that took the vaccine did in fact get covid-19, while 67% of the LGBTQ taking the placebo contracted covid-19.  A substantially lower percentage of people who took tha vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for the LGBTQ. The odds of getting covid-19 while taking the vaccine is 1:3. The LGBTQ that took the drug were half as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of LGBTQ that got covid-19 or not.
In the second plot, the one above, The left two bars of the graph are all the LGBTQ that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of LGBTQ who got covid-19, now represents the percentage of LGBTQ that got covid-19 after taking the placebo or taking the Vaccine. 


The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected, contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


## Numerical Results




```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of LGBTQ that got covid-19 after taking the placebo and taking the vaccine. 
Among the LGBTQ, 71% got covid-19 after taking the vaccine, while the approximately 29% of LGBTQ that were given the placebo contracted covid-19. The vaccine is clearly ineffective in preventing covid-19 in LGBTQ and it is not recommended. Those who took the placebo were more likely to contract covid-19.
The odds of getting covid-19 while taking the vaccine is 3:1 for LGBTQ. The LGBTQ that took the drug were much more likely to get covid-19 (71%) than those taking the placebo (29%).

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```

The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at .00006. This is an awfully small number. seeing this number means that these results are not likely to occur by chance.



The null hypothesis predicted that the civid-19 vaccine effectiveness is independent of subgroup. Since it is clear that subgroup does play a role in the rate of infection in individuals who were vaccinated, the alternative hypothesis, that there is a correlation between covid-19 infection rate in those vaccinated and subgroups, is the favorable option.

The odd ratio above illustrates a ratio of 0.4. The LGBTQ who received the vaccine were as likely to get covid-19 than those who received the placebo. 

```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```

```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for LGBTQ. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting LGBTQ from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the LGBTQ can be seen above as 29. The effectiveness confidence interval for the LGBTQ is 24.45 to 36.56.

# Conclusions and recommendations
We reject the null hypothesis, because there is a difference in covid-19 vaccine effectiveness among different subgroups (males,females, drug users and LGBTQ). Vaccine effectiveness varies between subgroups and each subgroup had a different P value, thus their different rate of effectiveness in all subgroups after receiving the vaccine.

Males vaccinated had 2:1 of contracting covid-19 after vaccination, thus it is ineffective for males with a P value of 0.0005801, which is too low. this indicates strong evidence against the null hypothesis, so we reject the null hypothesis. In conclusion males should not take the vaccine, because it is ineffective at protecting males from covid-19.

Females were 1:2 odds of contracting covid-19 after vaccination with a p value of <2e-16 (0.000000000000002) which is too low, this indicates strong evidence against the null hypothesis, so we reject the null hypothesis. In conclusion women should take the vaccine as it is effective at protecting females from covid-19.

Drug users were 1:2 odds of contracting covid-19 after vaccination with a p value of 0.002 which is too low, this indicates strong evidence against the null hypothesis, so we reject the null hypothesis. In conclusion drug users should take the vaccine as it is effective at protecting drug users from covid-19.

LGBTQ had a 5:3 odds of contracting covid-19 after vaccination with a p value of 0.00006 which is too low, this indicates strong evidence against the null hypothesis, so we reject the null hypothesis. In conclusion LGBTQ should not get the vaccine as it is ineffective at protecting females from covid-19.

To reiterate it is clear thatbthere is a difference in covid-19 vaccine effectiveness between different subgroups, thus we reject the null hypothesis. The vaccine is effective at fighting covid-19 for females and drug users, but ineffective for males and LGBTQ. 
